---
title: "MC5 Hobo Pressure Workflow"
output: html_document
---
The markdown file documents the data processing of Hobo pressure
records collected at the MC5 / McDowell Trib monitoring location.

***
# Standard Initialization and Needed Libraries
```{r}
## Clear stuff
rm(list=ls())
graphics.off()

library("zoo")
library("broom")

## Set working directory
## setwd("c:/Users/95218/Documents/R/MC5")
```

***
# Define a Function to Read In Raw .csv Files
The atmospheric and water pressure files are not necessarily
synchronized in time.  Use a spline fit to predict data on the 0:00,
0:05, 0:10 stamps, from the raw data which may be 0:02, 0:07, 0:12, ....

```{R}
## Define a function to run through a list of dfs, name cols, covert dates
readHoboInterp <- function(dataFile) {
    ## Read in file, rename resulting df columns
    data <- read.csv(paste0("c:/Users/95218/Documents/R/MC5/", dataFile),
                     sep=",", stringsAsFactors=FALSE, skip=1, header=TRUE)
    data <- data[ ,c(2,3,4)]
    names(data) <- c("dt", "hoboPSI", "hoboF")

    ## Format dates from excel files
    data$dt <- strptime(data$dt, format="%m/%d/%y %I:%M:%S %p",
                        tz="America/Panama")
    data$dt <- as.POSIXct(data$dt)

    ## Create regulalry spaced 5 minute time series overlapping data file
    regTS <- data.frame(dt=seq.POSIXt(to=tail(data$dt,6)[6], by=300,
                      from=round(data$dt[1],"mins")-
                          (round(data$dt[1],"mins")$min%%5)*60),kp=1)

    ## Merge even spaced dt with raw data frame
    dataMerged <- merge(data, regTS, by="dt", all=TRUE)

    ## Interpolate the psi series to the even 5 min time stamps
    zooObj <- zoo(dataMerged[2], dataMerged$dt)
    spline1 <- na.spline(zooObj, na.rm=FALSE)
    dataMerged$MC5ModlPsi <- coredata(spline1)

    ## Interpolate the temperature series to the even 5 min time stamps
    zooObj2 <- zoo(dataMerged[3], dataMerged$dt)
    spline2 <- na.spline(zooObj2, na.rm=FALSE)
    dataMerged$MC5ModlTempF <- coredata(spline2)

    ## drop all but dt, interp temp, interp psi
    dataMerged <- dataMerged[which(dataMerged$kp==1),c(1,5,6)]

    return(dataMerged)
}
```

# Create file lists of all air and water files in the current file directory

Code will do a recursive pattern match to find all files.  Note that
any .cvs contiaing "air" or "water" in file names will be pulled in.
```{r}
## Read the directories of files that CB sent
dataAir <- as.list(list.files(
    path="c:/Users/95218/Documents/R/MC5",
    pattern="Air.*\\.csv", recursive=TRUE))
names(dataAir) <- paste0("file", seq(1:length(dataAir)))

dataWater <- as.list(list.files(
    path="c:/Users/95218/Documents/R/MC5",
    pattern="Water.*\\.csv", recursive=TRUE))
names(dataAir) <- paste0("file",seq(1:length(dataAir)))
```

# Pass those file lists into the function readHoboInterp (defined above)
This will take the lists of file names generated in previous step and
send them through lapply() into the user defined function.  Then
rbind() the resulting lists into data frames, and sort.  The next few
steps drop some structural artifacts, convert units, cleanup. The 'Gap
Fill' steps on airFrame insert a regulalry spaced 5 minute timestamp
for later gap backfilling via CLT Airport data.

```{r}
## Pass df of water press. file names to f(), rbind results, sort
waterList <- lapply(dataWater, readHoboInterp)
waterFrame <- do.call("rbind", waterList)
waterFrame <- waterFrame[order(waterFrame$dt), ]

## Drop the [:,1] structure
dimnames(waterFrame$MC5ModlPsi) <- NULL
dimnames(waterFrame$MC5ModlTempF) <- NULL
waterFrame$MC5ModlPsi <- waterFrame$MC5ModlPsi[,1]
waterFrame$MC5ModlTempF <- waterFrame$MC5ModlTempF[,1]

## Some unit conversions
waterFrame$MC5ModlHPa <- waterFrame$MC5ModlPsi*68.94757
waterFrame$MC5ModlTempC <- (waterFrame[,"MC5ModlTempF"]-32)/1.8

## Pass df of air press. file names to f(), rbind results, sort
airList <- lapply(dataAir, readHoboInterp)
airFrame <- do.call("rbind", airList)
airFrame <- airFrame[order(airFrame$dt), ]

## Fill gaps with NAs, tempertature regression and pressure convr. from CLT
gapFill <- data.frame(dt=seq.POSIXt(to=tail(airFrame$dt,6)[6], by=300,
                      from=airFrame$dt[1]))
airFrame <- merge(airFrame, gapFill, by="dt", all=TRUE)

## Drop the [:,1] structure
dimnames(airFrame$MC5ModlPsi) <- NULL
dimnames(airFrame$MC5ModlTempF) <- NULL
airFrame$MC5ModlPsi <- airFrame$MC5ModlPsi[,1]
airFrame$MC5ModlTempF <- airFrame$MC5ModlTempF[,1]

## Some unit conversions
airFrame$MC5ModlHPa <- airFrame$MC5ModlPsi*68.94757
airFrame$MC5ModlTempC <- (airFrame[,"MC5ModlTempF"]-32)/1.8

rm(dataAir, dataWater, readHoboInterp, gapFill, waterList, airList)
```

***
# Plot the 5 minute data (interpolated to even stamps)
This plot shows that we have temporal gaps in both records.  Water is
not correctable, but air can potentially be backfilled by atmospheric
pressure data from elsewhere (e.g, CLT).

```{r include = TRUE}
## Plot spline fits of raw pressure files
dateTicks <- seq.POSIXt(from=min(waterFrame$dt,airFrame$dt,na.rm=TRUE),
                      to=max(waterFrame$dt,airFrame$dt,na.rm=TRUE),
                      by="month")

##graphics.off()
##pdf(file="MC5 Pressure.pdf", width=10, height=7.5)
dev.new(width=10,height=7.5, xpos=1930,ypos=65)
par(xaxs="i", yaxs="i", mai=c(1,1.5,.5,.5), font=2,
    cex.axis=1.2, family="serif", omi=rep(0,4))
plot(waterFrame$dt, waterFrame$MC5ModlPsi, pch=16,
     axes=FALSE, xlab="", ylab="", ylim=c(13,18), cex=.2)
points(airFrame$dt, airFrame$MC5ModlPsi, pch=16, col="red", cex=.2)
box()
axis.POSIXct(side=1, at=dateTicks, x=dateTicks, format="%b-%y")
axis(2, las=2)
mtext(side=2, "Pressure (psi)", line=3.5, cex=1.5, font=2)
legend("topleft", legend=c("Water", "Air"), pch=16, pt.cex=.75,
       col=c("black", "red"), bty="n", cex=1.5)
##dev.off()
```

***
# Load CLT Airport Data File
From [NOAA, Mecklenburg/CLT](https://www.ncdc.noaa.gov/cdo-web/datatools/lcd),
get CLT Airport climate data, save at a .csv.

```{r}
## Load in the Airport data, format and clean up NAs
cDoug <- read.csv("CLTMet.csv", header=TRUE, sep=",",
                  stringsAsFactors=FALSE)
cDoug <- cDoug[ ,c("DATE","HOURLYSeaLevelPressure",
                   "HOURLYStationPressure", "HOURLYDRYBULBTEMPC")]
cDoug$dt <- as.POSIXct(cDoug$DATE, format="%m/%d/%Y %H:%M",
                       tz="America/Panama")
## Rename, Slp=sea level press., Stp=station pressure.
names(cDoug) <- c("date", "CLTSlpHgIn", "CLTStpHgIn", "CLTTempC", "dt")

cDoug$CLTSlpHgIn <- as.numeric(cDoug$CLTSlpHgIn)
cDoug$CLTStpHgIn <- as.numeric(cDoug$CLTStpHgIn)
cDoug$CLTSlpHPa <- cDoug$CLTSlpHgIn*33.8639

cDoug$CLTTempC <- as.numeric(cDoug$CLTTempC)

```

# Plot the CLT pressure series
```{r}
## Plot CLT data
dateTicks <- seq.POSIXt(from=min(cDoug$dt, na.rm=TRUE),
                      to=max(cDoug$dt, na.rm=TRUE),
                      by="month")

##graphics.off()
##pdf(file="MC5 Pressure.pdf", width=10, height=7.5)
dev.new(width=10,height=7.5, xpos=1930,ypos=65)
par(xaxs="i", yaxs="i", mai=c(1,1.5,.5,.5), font=2,
    cex.axis=1.2, family="serif", omi=rep(0,4))
plot(cDoug$dt, cDoug$CLTSlpHgIn, pch=16,
     axes=FALSE, xlab="", ylab="", ylim=c(28,32), cex=.2)
points(cDoug$dt, cDoug$CLTStpHgIn, pch=16, col="red", cex=.2)
box()
axis.POSIXct(side=1, at=dateTicks, x=dateTicks, format="%b-%y")
axis(2, las=2)
mtext(side=2, "Pressure (HgIn)", line=3.5, cex=1.5, font=2)
legend("topleft", legend=c("Sea Level Pressure", "Station Pressure"),
       pch=16, pt.cex=.75, col=c("black", "red"), bty="n", cex=1.5)
##dev.off()
```
